Practical Machine Learning: Assignment Writeup
========================================================
##Executive summary

 In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 
 
The goal of your project is to predict the manner in which they did the exercise

#Data Preparation
```{r,echo=TRUE}
#Data loading
ppmlTrain=read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
ppmlTest=read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))
```

-all variables with at least one "NA" were excluded from the analysis. 
-Variables related to time and user information were excluded 
-Same variables were mainteined in the test data set to be used for predicting the 20 test cases

```{r,echo=TRUE}
## NA exclusion for all the variables
NONAppmlTrain=ppmlTrain[, apply(ppmlTrain, 2, function(x) !any(is.na(x)))] 
```

```{r,echo=TRUE}
## variables with user information, time and undefined
cleanedppmlTrain=NONAppmlTrain[,-c(1:8)]
```
```{r,echo=TRUE}
#20 test cases
cleanedppmltest=ppmlTest[,names(cleanedppmlTrain[,-52])]
```


#Partition and predicting

```{r,echo=TRUE}
library(randomForest)
library(caret)
Train=createDataPartition(y=cleanedppmlTrain$classe, p=0.75,list=F)
training=cleanedppmlTrain[Train,] 
test=cleanedppmlTrain[-Train,] 
```
#Conclusion

Random forest trees were generated for the training dataset using cross-validation. The algorithm was examnined with the partitioned test set to determine the accuracy and estimated error of prediction. 

```{r,echo=TRUE}
library(e1071)
library(randomForest)
library(caret)

fc=trainControl(method="cv", number=5, allowParallel=T, verbose=T)
rffit=train(classe~.,data=training, method="rf", trControl=fc, verbose=F)
```


```{r,echo=TRUE}
predictrf=predict(rffit, newdata=test)
confusionMatrix(predictrf, test$classe)
```

```{r,echo=TRUE}
predof20=predict(rffit, newdata=cleanedppmltest)
# prediction of the 20 cases
predof20

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predof20)
```
